Pesticide exposure in small streams in Germany

### 2016-09-13 article.tex
fixed model description

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title[Pesticides small streams]{Pesticides in small streams in Germany}
% \abbreviations{mo, neon, ra, tu, fw}
\keywords{Monitoring, Neonicotinoid, Risk Assessment Toxic Units, Freshwater}
% RAC,

@@

In this study we try to fill this gap and analyze nation-wide chemical monitoring data from SWB from Germany.
Using this large scale dataset we aimed to answer the following research questions:

\begin{enumerate}[(i)]
  \item Is currently available monitoring data suitable for a description of pesticide pollution?
	\item How does the relationship between agricultural land use and catchment size with pesticide pollution look like?
  Are there thresholds for where pesticide pollution becomes apparent?
  \item Are samples taken during or shortly after heavy rainfall events or different seasons more likely to indicate chemical pollution?
	\item  How polluted are SWB in Germany and which pesticides currently pose the biggest threat to them?
\end{enumerate}

Answering these questions might give important insights for future monitoring planning, ecological risk assessment and water resource management.

@@

\subsection{Characterization of catchments}
We compiled a total of 3,049 sampling sites with pesticide measurements.
We delineated catchments upstream for each of the sampling sites using a digital elevation model (DEM) \citep{eea_digital_2013} and the multiple flow direction algorithm \citep{holmgren_multiple_1994} as implemented in GRASS GIS 7 \citep{neteler_grass_2012}.
Catchment delineation was manually checked for accuracy by comparison with a stream network provided by the government.
The delineation algorithm produced only for 30\% of the sites accurate results.
For the rest we were able to compile catchment size data from authorities (47\% of sites) or drainage basins per stream segment provided by authorities (13\% of sites).
For 10\% of the sites we were not able to compile catchment size data.
For each derived catchment (either from DEM or drainage basins) we calculated the relative cover (in \%) with agricultural areas based on Official Topographical Cartographic Information System (ATKIS) of the land survey authorities \citep{adv_atkis_2016}.
We additionally used agricultural cover data provided by authorities (18\% of sites), which resulted to 21\% of sites with missing agricultural cover data. 
For 78\% of the sites both, the proportion of agricultural landuse and catchment size were available.
% see do_overview.R for numbers

@@

where $No_i$ is the observed number of exceedances at site $i$. 
We modeled $No_i$ as resulting from a negative binomial distribution ($NB$).
The proportion of agriculture within the catchment ($Agri_i$) and the catchment size of the site ($Size_i$) were used as predictors. 
$f_1$ and $f_2$ are smoothing functions using penalized cubic regression splines \citep{wood_generalized_2006}.
The degree of smoothness was estimated using restricted maximum likelihood (REML) during model fitting process \citep{wood_fast_2011}.
The number of sampling events per site ($n_i$) was used as an offset to account different sampling efforts at a site and is equivalent to modeling the rate of exceedances. 
We used point-wise 95\% Confidence Intervals (CI) of the first derivative of the fitted smooth to check if there are regions of statistically significant changes.
GAMs were fitted using the mgcv package \citep{wood_fast_2011}.

\todo{Überleitung}
RQ showed a skewed distribution with an excess of zeros (no pesticides with RAC detected). 
Therefore, we modeled these as two processes (one generating values below the limit of quantification (LOQ) and one generating values above LOQ) using a Zero-Adjusted Gamma (ZAGA) distribution \todo{Cite rigby 2010}:

\begin{align}
y \sim ZAGA(\mu, \pi) = 
  \begin{cases}
    (1 - \pi)   & \quad  \text{if } y < LOQ \\
    \pi \times f_{Gamma} (\mu) & \quad \text{if } y > LOQ \\
  \end{cases}
  \label{eqn:eqn3}
\end{align}

$\pi$ denotes the probability of a observation being above LOQ and $f_{Gamma}$ denotes the gamma function and is used for values greater LOQ.
We used the precipitation at sampling date ($prec_0$) and the day before ($prec_{-1}$), as well as quarters of the year ($season_{Q1-Q4}$) as linear predictors for the mean RQ ($\mu$) and the $\pi$:

\begin{align}
\begin{split}
\log(\mu) = \beta_0 + \beta_1 \times prec_0 + \beta_2 \times prec_{-1} + \beta_{3-5} \times season_{Q2-Q4} \\
logit(\pi) = \beta_0 + \beta_1 \times prec_0 + \beta_2 \times prec_{-1} + \beta_{3-5} \times season_{Q2-Q4} \\
\end{split}
\label{eqn:eqn4}
\end{align}

Changes in $\mu$ can be interpreted as changes in the absolute value of RQ, whereas changes in $\pi$ can be interpreted as changes in the probability that there is any risk.
To account for temporal autocorrelation and differences between federal states we used the site nested within state as random factors.
We fitted this model separatedly to each compound with a RAC, with at least 500 samples and more then 5\% of values above quantification limit.

\todo{check formulas, add i, add random effect? }

\todo{1 - remove columns from appendix table, 
2- keep only those with RACs,
3- indicate which where used for modeling here
}

@@

In total 484 different compounds used as pesticides and their metabolites were measured at least once (Supplement, Table S2 \todo{remove LC50 and MAC-EQS from table!}). 
Most of the compounds were herbicides (179), followed by insecticides (117) and fungicides (109).
Most samples were take in the months April till October, with less samples in the winter (see Supplemental Figure S2).
Only 5.5\% (160,800) of all measurements were detects above the limit of quantification (LOQ).
We found substantial differences in the spectra of analyzed compounds between federal states (Figure \ref{fig:fig2}).
Hierarchical clustering revealed three groups (see also Supplement Figure S2):

@@

The spatio-temporal intersection revealed that 5\% of the samples were taken at or after days with rainfall events greater than 10mm / day (Supplement, Figure S3).

\begin{figure}[ht]
  \includegraphics[width=0.95\textwidth]{figure5.pdf}
  \caption{Estimated coefficients and their 95\% CI for the model describe in equations \ref{eqn:eqn3} and \ref{eqn:eqn4}. Left column: Effect on mean RQ; Right column: effect on the probability that RQ \textgreater 0.
  Coefficient where the CI does not encompasses zero are shown in black.
  }
  \label{fig:fig5}
\end{figure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title[Pesticides small streams]{Pesticide exposure in small streams in Germany}
% \abbreviations{mo, neon, ra, tu, fw}
\keywords{Monitoring, Neonicotinoid, Risk Assessment Toxic Units, Freshwater}
% RAC,

@@

In this study we try to fill this gap and analyze large scale nation-wide chemical monitoring data from Germany.
First, we revise the available monitoring data if it is suitable for a large scale description of pesticide pollution.
Then we analyse the relationship between agricultural land use and catchment size with pesticide pollution and investigate potential thresholds.
We also investigate whether samples taken during or shortly after heavy rainfall events or different seasons are more likely to indicate chemical pollution.
Finally, we give an overview on the pollution of SWB in Germany.

% Answering these questions might give important insights for future monitoring planning, ecological risk assessment and water resource management.

@@

\subsection{Characterization of catchments}
We compiled a total of 3,049 sampling sites with pesticide measurements.
We delineated catchments upstream for each of the sampling sites using a digital elevation model (DEM) \citep{eea_digital_2013} and the multiple flow direction algorithm \citep{holmgren_multiple_1994} as implemented in GRASS GIS 7 \citep{neteler_grass_2012}.
Catchment delineation was manually checked for accuracy by comparison with a stream network provided by the government.
The delineation algorithm produced only for 30\% of the sites accurate results.
For the rest we were able to compile catchment size data from authorities (47\% of sites) or drainage basins per stream segment provided by authorities (13\% of sites).
For 10\% of the sites we were not able to compile catchment size data.
For each derived catchment (either from DEM or drainage basins) we calculated the relative cover (in \%) with agricultural areas based on Official Topographical Cartographic Information System (ATKIS) of the land survey authorities \citep{adv_atkis_2016}.
We additionally used agricultural cover data provided by authorities (18\% of sites), which resulted to 21\% of sites with missing agricultural cover data. 
For 78\% of the sites both, the proportion of agricultural land use and catchment size were available.
% see do_overview.R for numbers

@@

where $No_i$ is the observed number of exceedances at site $i$. 
We modeled $No_i$ as resulting from a negative binomial distribution ($NB$).
The proportion of agriculture within the catchment ($Agri_i$) and the catchment size of the site ($Size_i$) were used as predictors. 
$f_1$ and $f_2$ are smoothing functions using penalized cubic regression splines \citep{wood_generalized_2006}.
The degree of smoothness was estimated using restricted maximum likelihood (REML) during model fitting process \citep{wood_fast_2011}.
The number of samples per site ($n_i$) was used as an offset to account different sampling efforts (sampling interval and analysed compound spectrum) at a site and is equivalent to modeling the rate of exceedances. 
We used point-wise 95\% Confidence Intervals (CI) of the first derivative of the fitted smooth to check if there are regions of statistically significant changes.
GAMs were fitted using the mgcv package \citep{wood_fast_2011}.

While agricultural land use and catchment size vary only between sites, this is not the case for precipitation that changes also with time.
Therefore, we modeled the effects of precipitation in a separate model.
RQ showed a skewed distribution with an excess of zeros (no pesticides with RAC detected). 
Therefore, we modeled these as two processes (one generating values below the limit of quantification (LOQ) and one generating values above LOQ) using a Zero-Adjusted Gamma (ZAGA) distribution \todo{Cite rigby 2010}:

\begin{align}
RQ_i \sim ZAGA(\mu_i, \pi_i) = 
  \begin{cases}
    (1 - \pi_i)   & \quad  \text{if } y < LOQ \\
    \pi_i \times f_{Gamma} (\mu_i) & \quad \text{if } y \ge LOQ \\
  \end{cases}
  \label{eqn:eqn3}
\end{align}

$\pi_i$ denotes the probability of a observation i being above LOQ and $f_{Gamma}$ denotes the gamma function and is used for values greater LOQ, where $\mu$ is the mean of the gamma function.
We used the precipitation at sampling date ($prec_0$) and the day before ($prec_{-1}$), as well as quarters of the year ($season_{Q1-Q4}$) as linear predictors for $\mu$ and $\pi$.
To account for temporal autocorrelation and differences between federal states we used the site nested within state as random intercepts:

\begin{align}
\begin{split}
\log(\mu_i) = prec_{0 ijk} + prec_{-1 ijk} + season_{ijk} + state_j + site_{jk}\\
logit(\pi_{ijk}) = prec_{0 ijk} + prec_{-1 ijk} + season_{ijk} + state_j + site_{jk}\\
state_{j} \sim N(0, \sigma^2_{state}) \\
site_{jk} \sim N(0, \sigma^2_{site})
\end{split}
\label{eqn:eqn4}
\end{align}

Changes in $\mu$ can be interpreted as changes in the absolute value of RQ, whereas changes in $\pi$ can be interpreted as changes in the probability that there is any risk.
We fitted this model separately to each compound with a RAC, with at least 1000 samples and more then 5\% of values above quantification limit (see Supplement, Table S3 for a list of compounds). 
We implemented the model using the gamlss packages \cite{}

@@

In total 484 different compounds used as pesticides and their metabolites were measured at least once (Supplement, Table S2). 
Most of the compounds were herbicides (179), followed by insecticides (117) and fungicides (109).
Most samples were take in the months April till October, with less samples in the winter (see Supplemental Figure S2).
Only 5.5\% (160,800) of all measurements were detects above the limit of quantification (LOQ).
We found substantial differences in the spectra of analyzed compounds between federal states (Figure \ref{fig:fig2}).
Hierarchical clustering revealed three groups (see also Supplement Figure S2):

@@

The spatio-temporal intersection revealed that 5\% of the samples were taken at or after days with rainfall events greater than 10mm / day (Supplement, Figure S3).
Precipitation at the day before sampling increases $\mu$ and also $\pi$ for nearly all modeled compounds. 
The effects were less pronounced ($\pi$) or not directed ($\mu$) for precipitation at the day of sampling (Figure \ref{fig:fig5}, top).

Quarters 2-4 showed higher RQ than in Q1, except for the compounds  (Figure \ref{fig:fig5}, bottom)

\begin{figure}[ht]
  \includegraphics[width=0.95\textwidth]{figure5.pdf}
  \caption{Estimated coefficients and their 95\% CI for the model describe in equations \ref{eqn:eqn3} and \ref{eqn:eqn4} for each of the 24 analysed compounds (see Supplement Table . Left column: Effect on mean RQ; Right column: effect on the probability that RQ \textgreater 0.
  Coefficients where the CI encompasses zero are shown in grey.
  }
  \label{fig:fig5}
\end{figure}
